---
title: "RWorksheet_Magallanes#1"
author: "Killy Magallanes"
date: "2024-09-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1.Set up a vector named age.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29,35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42,53, 41,51, 35, 24, 33, 41)
```
a.How many data points.
```{r}
num_data_points <- length(age)
num_data_points
```
2. Find the Reciprocal of the values of age.
```{r}
reciprocal_age <- 1 / age
reciprocal_age
```
3. Assign also new_age <- c(age, 0, age).
```{r}
new_age <- c(age, 0, age)
new_age
```
what  happened to  the  new_age?
- The new_age has 0 between the 2 ages

4. Sort the values for age.
```{r}
sort(age)
```

5. Find the minimum and maximum value for age.
```{r}
min(age)
max(age)
```

6. Set up a vector named data.
```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
```
How many data points?
```{r}
nmbrofdtps <- length(data)
nmbrofdtps
```


7. Generate a new vector for data where you double every value ofthe data.
```{r}
newData <- 2 * data
newData
```
what happened to  the data?
-The data doubles and was multiplied by  2

8. Generate a sequence for the following scenario.

8.1 Integers from 1 to 100.
```{r}
Int <- seq(1:100)
Int
```
8.2 Numbers from 20 to 60.
```{r}
Num <- seq(from=20,to=60)
Num
```
8.3 Mean of numbers from 20 to 60.
```{r}
Num <- seq(from=20, to=60)
num20to60 <- mean(Num)
num20to60
```
8.4 Sum of numbers from 51 to 91.
```{r}
Num <- sum(51:91)
Num
```
8.5 Integers from 1 to 1000.
```{r}
Int <- seq(1:1000)
Int
```
a. How many data points from 8.1 to 8.4?
```{r}
datapoints <- length(8.1:8.4)
datapoints
```
c. For 8.5, find only the maximum datapoints until 10
```{r}
mxdtpts <- Int[1:10]
mxdtpts
length(mxdtpts)
```
9. Print a vector with the integers between 1 and 100 that are notdivisible by 3, 5 and 7 using filte roption
```{r}
Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(100))
```
10. Generate a sequence backwards of the integers from 1 to 100.
```{r}
rev(seq(1:100))
```
11. List all the natural numbers below 25 that are multiples of 3or 5.
```{r}
x <- Filter(function(i) {any(i %% c(3,5) ==0)}, seq(24))
Sum_of_x <- sum(x)
Sum_of_x
```

a. How many data points from 10 to 11?
```{r}
datapt <- length(10:11)
datapt
```
12.
```{r}
#x <- {0 + x + 5 + }
#This string has syntax error therefore it could'nt run.
```
13.
```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75, 75,77.)
s2 <- score[2]
s2
s3 <- score[3]
s3
```
14. *Create a vector a = c(1,2,NA,4,NA,6,7).
```{r}
a <- c(1, 2 ,NA, 4, NA, 6, 7)
print(a,na.print="-999")
print(a,na.print="-999")
```
15
```{r}
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
(R.version.string)
```




