---
title: "RWorksheet_Magallanes#4b"
author: "Killy Magallanes"
date: "2024-10-28"
output: pdf_document
---

Using Loop Function
for() loop

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrix_result <- matrix(0, nrow = 5, ncol = 5)
for (i in 1:5) {
  for (j in 1:5) {
    matrix_result[i, j] <- abs(vectorA[i] - vectorA[j])
  }
}
print(matrix_result)
```
2. Print the string "*" using for() function. The output should be the same as shown
in Figure

```{r}
num_lines <- 5

for (i in 1:num_lines) {
  cat(rep("*", i), "\n")
}
```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r}
start_num <- as.numeric(readline(prompt = "Enter a number to start the Fibonacci sequence: "))

a <- 0
b <- 1

repeat {
  next_fib <- a + b
  a <- b
  b <- next_fib
  if (next_fib > 500) {
    break
  }
  if (next_fib >= start_num) {
    cat(next_fib, "\n")
  }
}
```
Using Basic Graphics (plot(),barplot(),pie(),hist())
4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result

```{r setup, include=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org/"))
install.packages("readxl")
library(readxl)
ShoeSizes <- read_excel("C:\\Users\\killy\\Documents\\pushed\\RWorksheet_Magallanes#4b\\household_Data.xlsx")
ShoeSizes
```

b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.

```{r}
female_subset <- subset(ShoeSizes, Gender == "F")
male_subset <- subset(ShoeSizes, Gender == "M")
num_female <- nrow(female_subset)
num_male <- nrow(male_subset)
cat("Number of Female observations:", num_female, "\n")
cat("Number of Male observations:", num_male, "\n")

```

c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.

```{r}
gender_count <- table(ShoeSizes$Gender)
bar_colors <- c("blue", "purple")
barplot(gender_count, 
        main = "Number of Males and Females in Household Data",
        xlab = "Gender",
        ylab = "Number of Observations",
        col = bar_colors,
        names.arg = c("Female", "Male"),
        legend.text = TRUE) 
```

5. 
a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.

```{r}
categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
values <- c(60, 10, 5, 25)
percentages <- round(values / sum(values) * 100)
pie_colors <- c("blue", "purple", "yellow", "black")
pie(percentages,
    labels = paste(categories, percentages, "%"),
    col = pie_colors, 
    main = "Expenses Breakdown", 
    radius = 1)
```
6. Use the iris dataset
a. a. Check for the structure of the dataset using the str() function. Describe what you
have seen in the output.

```{r}
data(iris)
str(iris)
```

b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?

```{r}
data(iris)
means <- colMeans(iris[, 1:4])
means
```
c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write
the R script and its result.

```{r}
data(iris)
species_count <- table(iris$Species)
pie_colors <- c("lightgreen", "lightblue", "lightcoral")
pie(species_count,
    labels = paste(names(species_count), "\n", species_count),
    col = pie_colors,
    main = "Species Distribution in Iris Dataset", 
    radius = 1)  
```
d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.
```{r}
data(iris)

setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")

last_six_setosa <- tail(setosa, 6)
last_six_versicolor <- tail(versicolor, 6)
last_six_virginica <- tail(virginica, 6)

cat("Last six rows of Setosa:\n")
print(last_six_setosa)
cat("\nLast six rows of Versicolor:\n")
print(last_six_versicolor)
cat("\nLast six rows of Virginica:\n")
print(last_six_virginica)
```

e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal
width and length, labels for the x and y axis, the pch symbol and colors should be based on
the species.

Hint: Need to convert to factors the species to store categorical variables.

```{r}
data(iris)

iris$Species <- as.factor(iris$Species)

species_colors <- c("setosa" = "lightgreen", "versicolor" = "lightblue", "virginica" = "lightcoral")


plot(iris$Sepal.Length, iris$Sepal.Width, 
     col = species_colors[iris$Species], 
     pch = 19, 
     main = "Iris Dataset", 
     sub = "Sepal Width and Length",
     xlab = "Sepal Length",  
     ylab = "Sepal Width") 

legend("topright", legend = levels(iris$Species), 
       col = species_colors, pch = 19)
```

f. Interpret the results

The scatterplot of Sepal.Length and Sepal.Width from the iris dataset reveals distinct differences among the three species. Setosa (light green) is clearly separated from Versicolor (light blue) and Virginica (light coral), exhibiting smaller sepal dimensions, which allows for easy identification. Versicolor shows a moderate range of sepal lengths and widths, while Virginica occupies the upper right area of the plot, indicating it has the largest sepals. There is some overlap between Versicolor and Virginica, suggesting that they may be more challenging to distinguish based solely on these measurements. Overall, the plot highlights a positive correlation between sepal length and width for Versicolor and Virginica, providing valuable insights for species classification in botanical research and data analysis.

7. 
```{r}
Alexafile <- read_excel("C:\\Users\\killy\\Documents\\pushed\\RWorksheet_Magallanes#4b\\alexa_file.xlsx")
Alexafile
```
a. Rename the white and black variants by using gsub() function.

```{r}
RObject$columnName <- gsub("Old Name", "New Name", RObject$columnName)
```

