---
title: "RWorksheet_Magallanes#4a"
author: "Killy Magallanes"
date: "2024-10-14"
output: pdf_document
---

1. The table below shows the data about shoe size and height. Create a data frame. 


```{r}
Household_Data <- read_excel("C:\\Users\\killy\\Documents\\pushed\\RWorksheet_Magallanes#4a\\household_Data.xlsx")
head(Household_Data, n = 10)
```
a. It displays the shoe size, height, and gender in column

b. Create a subset by males and females with their corresponding shoe size and height. What is the result? show the Rscript
```{r}
female_data <- subset(Household_Data, Gender == "F")

male_data <- subset(Household_Data, Gender == "M")

female_data
male_data
```

c. Find the mean of shoe size  and height of the respondents. write the R scripts  and its result.

```{r}
mean_shoesize <- mean(Household_Data$Shoe_size)
mean_shoesize

mean_height <- mean(Household_Data$Height)
mean_height
```
d. Is there a relationship between shoe size and height? Why?

- Yes, the taller the respondent, the longer his/her feet

FACTORS

2. construct character vectors months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below thee actual values.

Consider data consisting of the names of months:
"March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August", "July","December","August","August","September","November","February",|
April")

```{r}
months_vector <- c("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")

factor_months_vector <- factor(months_vector)
factor_months_vector
print(levels(factor_months_vector))
```
3. Then check the summary() of the months_vector and factor_months_vector. Interpret the results of both vectors. Are they both equally useful in this  case?


```{r}
summary(months_vector)


summary(factor_months_vector)
```
Summary function will return basic information about the vector, such as the length (i.e., the total number of values) and other high-level characteristics. It returns a count of how many times each factor level appears in the vector.


4. Create a vector and factor for the table
```{r}
direction_vector <- c("East", "West", "North")
direction_vector
frequency_vector <- c(1,4,3)
frequency_vector

direction_factor <- factor(direction_vector)
direction_factor

frequency_factor <- factor(frequency_vector)
frequency_factor
```

5. Enter the data below in Excel with the file name = import_march.csv

a. Import the excel file into the Environment pane using read.table() function. Write the code.
```{r}
data <- read.table("C:/Users/killy/Documents/pushed/RWorksheet_Magallanes#4a/import_march.csv", header=TRUE, sep=",",stringsAsFactors = FALSE)
head(data)

```

b. View the dataset. Write the R scripts and its result

```{r}
View(data)
```





