---
title: "RWorksheet_Magallanes#4a"
author: "Killy Magallanes"
date: "2024-10-14"
output: pdf_document
---

1. The table below shows the data about shoe size and height. Create a data frame. 


```{r setup, include=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org/"))
install.packages("readxl")
library(readxl)
Household_Data <- read_excel("C:\\Users\\killy\\Documents\\pushed\\RWorksheet_Magallanes#4a\\household_Data.xlsx")
print(Household_Data)
```
a. It displays the shoe size, height, and gender in column

b. Create a subset by males and females with their corresponding shoe size and height. What is the result? show the Rscript
```{r}
female_data <- subset(Household_Data, Gender == "F")

male_data <- subset(Household_Data, Gender == "M")

female_data
male_data
```

c. Find the mean of shoe size  and height of the respondents. write the R scripts  and its result.

```{r}
mean_shoesize <- mean(Household_Data$Shoe_size)
mean_shoesize

mean_height <- mean(Household_Data$Height)
mean_height
```
d. Is there a relationship between shoe size and height? Why?

- Yes, the taller the respondent, the longer his/her feet

FACTORS

2. construct character vectors months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below thee actual values.

Consider data consisting of the names of months:
"March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August", "July","December","August","August","September","November","February",|
April")

```{r}
months_vector <- c("March", "April", "January", "November", "January", "September", "October", "September", "November", "August", "January", "November", "November", "February", "May", "August", "July", "December", "August", "August", "September", "November", "February", "April")

factor_months_vector <- factor(months_vector)
factor_months_vector
print(levels(factor_months_vector))
```
3. Then check the summary() of the months_vector and factor_months_vector. Interpret the results of both vectors. Are they both equally useful in this  case?


```{r}
summary(months_vector)


summary(factor_months_vector)
```
Summary function will return basic information about the vector, such as the length (i.e., the total number of values) and other high-level characteristics. It returns a count of how many times each factor level appears in the vector.


4. Create a vector and factor for the table
```{r}
direction_vector <- c("East", "West", "North")
direction_vector
frequency_vector <- c(1,4,3)
frequency_vector

direction_factor <- factor(direction_vector)
direction_factor

frequency_factor <- factor(frequency_vector)
frequency_factor
```

5. Enter the data below in Excel with the file name = import_march.csv

a. Import the excel file into the Environment pane using read.table() function. Write the code.
```{r}
data <- read.table("C:/Users/killy/Documents/pushed/RWorksheet_Magallanes#4a/import_march.csv", header=TRUE, sep=",",stringsAsFactors = FALSE)
head(data)

```

b. View the dataset. Write the R scripts and its result

```{r}
View(data)
```


Using Conditional Statements (IF-ELSE)
6. Full Search

a. Create an R Program that allows the User to randomly select numbers from 1 to 50.
Then display the chosen number. If the number is beyond the range of the selected choice,
it will have to display a string “The number selected is beyond the range of 1 to 50”. If
number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the
input number.

```{r}
number <- readline(prompt="Enter a number from 1-50")
number
```


```{r}
if (number <= 1 && number >=50){
  print("The number selected is beyond the range of 1 to 50")
} else if (number == 20){
  print("TRUE")
} else{
  print(number)
}
```
7. Change
a. 

```{r}
min_bills <- function(price) {
  bill_count <- 0
  
  if (price >= 1000) {
    bill_count <- bill_count + price %/% 1000
    price <- price %% 1000
  }
  
  if (price >= 500) {
    bill_count <- bill_count + price %/% 500
    price <- price %% 500
  }
  
  if (price >= 200) {
    bill_count <- bill_count + price %/% 200
    price <- price %% 200
  }
  
  if (price >= 100) {
    bill_count <- bill_count + price %/% 100
    price <- price %% 100
  }
  
  if (price >= 50) {
    bill_count <- bill_count + price %/% 50
  }
  
  print(bill_count)
}
min_bills(1250) 
```
8. 

a. Create a dataframe from the above table. Write the R codes and its output.
```{r}
Math_Score <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90))

Math_Score


```
b. Without using the rowMean function, output the average score of students whose average
math score over 90 points during the semester. write R code and its output.

```{r}
for (i in 1:nrow(Math_Score)) {
  avg_score <- sum(Math_Score[i, 2:5]) / 4
  print(paste0(Math_Score$Name[i], "'s average grade this semester is: ", avg_score))
}
```
c. Without using the mean function, output as follows for the tests in which the average
score was less than 80 out of 4 tests.
```{r}
for (j in 2:ncol(Math_Score)) {
  avg_test_score <- sum(Math_Score[, j]) / nrow(Math_Score)  
  if (avg_test_score < 80) {
    test_number <- j - 1
    print(paste("The", test_number, "th test was difficult."))
  }
}
```

d. Without using the max function, output as follows for students whose highest score for
a semester exceeds 90 points.
Example Output: Annie’s highest grade this semester is 95.

```{r}
for (i in 1:nrow(Math_Score)) {
  highest_score <- Math_Score[i, 2]  
  for (j in 3:ncol(Math_Score)) {
    if (Math_Score[i, j] > highest_score) {
      highest_score <- Math_Score[i, j]  # Update if a higher score is found
    }
  }
 if (highest_score > 90) {
    print(paste(Math_Score$Name[i], "'s highest grade this semester is", highest_score))
  }
}

```






